Adapting the MVDR to a GPU for active sonar systems of different sizes.

### 2013-01-22 ASA2013/asa2013.tex
la2pdf commit - 2013:01:22	20:44:26	(Tue22.Jan)

Suppose that a backscattered wavefield is recorded by an $M$ element uniform linear array, that a suitable set of delays are applied to the channels for the pixel of interest, and let the delayed output from the $m$'th channel be given as $x_m$. By definition a beamformer's output $z[n]$ at time instant $n$ is a weighted sum all the delayed data outputs:
\begin{align}
z[n] = \w\H[n]\x[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n]\\x_1[n]\\\vdots\\x_{M-1}[n]},\label{z}
\end{align}
where $w_m$ is the weight factor assigned to channel $m$. With static weights this would be referred to as the conventional delay-and-sum (DAS) beamformer. Various weighting patterns exists here for trading lateral resolution for improved noise suppression (contrast), but one always ends up with a compromise between the two\todo{don't quite like this one}~\cite{Harris1978}.

@@

\begin{figure}[t]\centering%
\graphicsAI[drawing,width=0.7\linewidth]{gfx/implementation.svg}%
\caption{MVDR beamforming. For one of the total number of pixels in range and azimuth, $N_y$ and $N_x$,\newline
1. an $L\times{}L$ sample covariance matrix $\eR$ is computed, \newline
2. the term $\eR^{-1}\1$ is found using a linear equation solver,\newline
3. and the beamformer output is computed from $z$ from (\ref{finalZ}), where $\w$ is found by substituting $\eR^{-1}\1$ into (\ref{weights}). }\label{mvdr_beamforming}
\end{figure}
The MVDR beamformer target this limitation by adapting the weights for each pixel to best fit the dynamic nature of the incoming wavefield. It computes the set of weights that minimizes the noise and signal power accumulated by the array, but with the constraint of unity gain for signal coming from the pixel of interest. When the data is delayed this is formulated as follows:
\begin{gather}
\argmin{\w}\ E\Big\{\big|\w\H[n]\x[n]\big|^2\Big\} \qquad\text{subject to}\qquad \w\H[n]\1 = 1. \label{mvdr_def}
% \vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},\label{weights}
\end{gather}
This is a convex optimization problem with the solution:
\begin{gather}
\vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},\label{weights}
\end{gather}
where $\1$ is a row vector of ones that represents broadside steering, and $\R=E\{\x[n]\x\H[n]\}$ is the spatial covariance matrix for the full array. Since $\R$ is unknown, we estimate it by by computing a sample covariance matrix $\eR$. In this computation we perform some degree of \emph{spatial averaging} to avoid signal cancellation, \emph{temporal averaging} to maintain true speckle statistics, and \emph{diagonal loading} to improve robustness to parameter errors ~\cite{Synnevag2009a}. Combined, these techniques will also ensure a numerically well conditioned $\eR$.

In terms of computational complexity the two important  of the MVDR method the 
Before If we let $x_l[n]$ represent the datavector from subarray $l$\todo{replace with with a drawing?},
\begin{gather}
\x_l[n] = \bmat{x_l[n] & x_{l+1}[n] & \dots & x_{l+L-1}[n]}\T,
\end{gather}
then a sample covariance matrix formed with temporal and spatial averaging, $\eR$, is given as:
\begin{gather}
\eR[n] =  \frac{1}{N_K N_L} \sumb{l=0}{M-L}\sumb{n'=n-K}{n+K} \x_l[n]\x_l\H[n] \in\mathbb{C}^{L,L},\label{spatialR}
\end{gather}
where $N_K = 2K+1$ is the number of temporal samples to perform averaging over, and $N_L = M-L+1$ is the number of subarrays. Finally, a fraction of the total power of $\eR[n]$ is added to its diagonal~\cite{Synnevag2007}.

>>>>

Suppose that a backscattered wavefield is recorded by an $M$ element uniform linear array, that a suitable set of delays are applied to the channels for the pixel of interest, and let the delayed output from the $m$'th channel be given as $x_m$. By definition a beamformer's output $z[n]$ at time instant $n$ is a weighted sum all the delayed data outputs:
\begin{align}
z[n] = \w\H[n]\x[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n]\\x_1[n]\\\vdots\\x_{M-1}[n]},\label{z}
\end{align}
where $w_m$ is the weight factor assigned to channel $m$. With static weights this would be referred to as the conventional delay-and-sum (DAS) beamformer. Various weighting patterns exists here for trading lateral resolution for improved noise suppression (contrast), but one always ends up with a compromise between the two\todo{don't quite like this one}~\cite{Harris1978}. However, there exists a class of adaptive beamformers that target this limitation by adapting the weights to the impinging wavefield. The MVDR method is one such technique.

@@

\begin{figure}[t]\centering%
\graphicsAI[drawing,width=0.7\linewidth]{gfx/implementation.svg}%
\caption{MVDR beamforming. For one of the total number of pixels in range and azimuth, $N_y$ and $N_x$,\newline
1. an $L\times{}L$ sample covariance matrix $\eR$ is computed, \newline
2. the term $\eR^{-1}\1$ is found using a linear equation solver,\newline
3. and the beamformer output is computed from $z$ from (\ref{finalZ}), where $\w$ is found by substituting $\eR^{-1}\1$ into (\ref{weights}). }\label{mvdr_beamforming}
\end{figure}
The basic idea behind the MVDR method is to compute the set of weights that minimizes the noise and signal power accumulated by the array, but with the constraint of unity gain for signals coming from the pixel of interest. When the data is delayed this can be formulated as follows:
\begin{gather}
\argmin{\w}\ E\Big\{\big|\w\H[n]\x[n]\big|^2\Big\} \qquad\text{subject to}\qquad \w\H[n]\1 = 1. \label{mvdr_def}
% \vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},\label{weights}
\end{gather}
This is a convex optimization problem with the solution:
\begin{gather}
\vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},\label{weights}
\end{gather}
where $\1$ is a row vector of ones that represents broadside steering, and $\R=E\{\x[n]\x\H[n]\}$ is the spatial covariance matrix for the full array. Since $\R$ is unknown, we estimate it by by computing a sample covariance matrix $\eR$. In this computation we perform some degree of \emph{spatial averaging} to avoid signal cancellation, \emph{temporal averaging} to maintain true speckle statistics, and \emph{diagonal loading} to improve robustness to parameter errors ~\cite{Synnevag2009a}. Combined, these techniques will also ensure a numerically well conditioned $\eR$.

In terms of computational complexity the diagonal loading factor is negligible, and will therefore be ignored from now on. Now let $x_l[n]$ represent the datavector from subarray $l$\todo{replace with with a drawing?},
\begin{gather}
\x_l[n] = \bmat{x_l[n] & x_{l+1}[n] & \dots & x_{l+L-1}[n]}\T,
\end{gather}
then a sample covariance matrix formed with temporal and spatial averaging, $\eR$, is given as:
\begin{gather}
\eR[n] =  \frac{1}{N_K N_L} \sumb{l=0}{M-L}\sumb{n'=n-K}{n+K} \x_l[n]\x_l\H[n] \in\mathbb{C}^{L,L},\label{spatialR}
\end{gather}
where $N_K = 2K+1$ is the number of temporal samples to perform averaging over, and $N_L = M-L+1$ is the number of subarrays. Finally, a fraction of the total power of $\eR[n]$ is added to its diagonal~\cite{Synnevag2007}.
